simulation_study_spline

### 2025-01-19 R/simulation_study_CRCVM/estimation_local/simulation_study_local.Rmd
update local simulation study

```{r}
library(here)
library(sf)
library(smoothSDE)
library(doParallel)
library(ggplot2)
library(dplyr)
library(mgcv)

@@

```{r hyperparameters}
N_ID=6
TMAX=1*24
SP_DF=c(7,4)
TAU_0=1.5
NU_0=4
DMIN=0.5
DELTA=1/360
PAR0_RACVM=c(0,0,1,1,0)
SIGMA_OBS_low=0.01
SIGMA_OBS_high=0.04
BY=30

@@

seed=100
set.seed(seed)
# Set the path to the directory containing the data
greenland_data_path <- here("Data","preprocessed_data","greenland")

@@

v0=c(0,0)

@@

i=1
while (i<=N_ID) {
  #choose location uniformly in the map
  x=c(runif(1,min=430,max=500),runif(1,min=7760,max=7900))
  if (!(is_in_land(st_point(x),border))) {
    p=nearest_boundary_point(st_point(x),border)
    Dshore=(x[1]-p[1])^2+(x[2]-p[2])^2
    #keep it as initial position if it is at least 50 metres away from the shore
    if (Dshore>DMIN && Dshore<1) {
      x0[i,]=x
      i=i+1
    }
  }
}

@@

fomega_cubic=function(cov_data,a=2,D0=0.5,D1=0.8,sigma_theta=pi/8,sigma_D=0.3,b=5){
  Dshore=cov_data$DistanceShore
  theta=cov_data$theta
  if (is.null(Dshore)){
    Dshore=1/cov_data$Ishore
  }

@@

@@

  res=sim_CRCVM(ftau=ftau_constant,fomega=fomega_cubic,fnu=fnu_constant,
                      log_sigma_obs=NULL,v0=v0,x0=x0[i,],times=times,land=border,verbose=FALSE)

  data_sim=res$sim
  data_sim$ID=factor(rep(i,length(data_sim$y1)))
  data_sim
}

@@

```{r}
# add noise
low_noise=rmvn(n_obs,rep(0,2),diag(rep(SIGMA_OBS_low^2,2)))
high_noise=rmvn(n_obs,rep(0,2),diag(rep(SIGMA_OBS_high^2,2)))
observed_data_le=data_sub
observed_data_le[,c("y1","y2")]=data_sub[,c("y1","y2")]+low_noise
observed_data_he=data_sub
observed_data_he[,c("y1","y2")]=data_sub[,c("y1","y2")]+high_noise

#subsample
data_lf_he=observed_data_he[seq(1,length(data_sub$time),by=BY),]
data_lf_le=observed_data_le[seq(1,length(data_sub$time),by=BY),]
data_hf_le=observed_data_le[seq(1,length(data_sub$time),by=6),]
data_hf_he=observed_data_he[seq(1,length(data_sub$time),by=6),]

@@

plot=ggplot()+geom_sf(data=border$geometry,fill="grey")+
  coord_sf(datum=st_crs("+init=EPSG:32626 +units=km"))+
  geom_point(data=data_lf_he,mapping=aes(y1,y2,col=ID),size=0.1)+
  geom_path(data=data_lf_he,mapping=aes(y1,y2,col=ID),size=0.1)+
  geom_point(data = data%>% filter(!duplicated(ID)),
             aes(x = y1, y = y2), shape = 3, size = 4, col = "red")+
  xlab("x") + ylab("y")

ggsave(filename=paste("plot_","border","_",TMAX,"h_",N_ID,"ID_",DMIN,"km",seed,".png",sep=""),plot=plot,width=10,height=5)

@@

add_covs=function(data) {
  #data has at least columns "time","y1" "y2", "p1","p2"
  #this function new data frame with three more columns "theta", "DistanceShore" and "ExpShore"

  new_data=data
  #observed deviation angles
  n=length(new_data$time)

  new_data$theta=rep(NA,n)

  for (id in unique(new_data$ID)) {

    #filter data with specific id
    sub_ind=(new_data$ID==id)
    sub_data=new_data[sub_ind,]
    n_sub=length(sub_data$time)

    #time steps
    dtimes=sub_data[2:n_sub,"time"]-sub_data[1:(n_sub-1),"time"]

    #step lengths
    dx=sub_data[2:n_sub,"y1"]-sub_data[1:(n_sub-1),"y1"]
    dy=sub_data[2:n_sub,"y2"]-sub_data[1:(n_sub-1),"y2"]

    #matrix of empirical velocity
    vexp_df=cbind(dx/dtimes,dy/dtimes)

    #nearest points on shore
    sub_data=cbind(sub_data,nearest_boundary_points(as.matrix(sub_data[,c("y1","y2")]),border))

    #matrix of normal vectors
    normal=as.matrix(sub_data[2:n_sub,c("y1","y2")]-sub_data[2:n_sub,c("p1","p2")])

    #angle between velocity and normal vector
    theta_coast=signed_angle(normal,vexp_df)

    #adjust lengths 
    theta_coast=c(theta_coast,1)

    new_data[sub_ind,"theta"]=theta_coast

    Dshore=rep(0,n_sub)
    for (i in 1:n_sub) {
      y1=sub_data[i,"y1"]
      y2=sub_data[i,"y2"]
      if (!(is_in_land(st_point(c(y1,y2)),border))) {
        p1=sub_data[i,"p1"]
        p2=sub_data[i,"p2"]
        Dshore[i]=sqrt((y1-p1)^2+(y2-p2)^2)
      }
    }
    new_data[sub_ind,"DistanceShore"]=Dshore

  return(new_data)
}

@@

```{r}
data_lf_he=add_covs(data_lf_he)
data_hf_le=add_covs(data_hf_le)
data_lf_le=add_covs(data_lf_le)
data_hf_he=add_covs(data_hf_he)
```

@@

ggplot(data_lf_he)+geom_histogram(aes(x=DistanceShore),bins=30)+facet_wrap(~ID)+theme_minimal()
quantile(data_lf$DistanceShore)
```

@@

ctcrw_lf_he<- SDE$new(formulas = formulas,data =data_lf_he[data_lf_he$ID %in% c(1,2,3),],type = "CTCRW",
                    response = c("y1","y2"),par0 = c(0,0,1,1),fixpar=c("mu1","mu2"),
                    other_data=list("log_sigma_obs0"=log(0.05)))

@@

ctcrw_lf_le<- SDE$new(formulas = formulas,data =data_lf_le[data_lf_le$ID %in% c(1,2,3),],type = "CTCRW",
                    response = c("y1","y2"),par0 = c(0,0,1,1),fixpar=c("mu1","mu2"),
                    other_data=list("log_sigma_obs0"=log(0.03)))

@@

#points on the surface perturbed by gaussian noise
f <- fomega_cubic(samples)
y <- f+0.1*rnorm(n)

@@

H=array(rep(SIGMA_OBS_high^2*diag(2),length(data_lf_he$time)),dim=c(2,2,length(data_lf_he$time)))

formulas <- list(mu1=~1,mu2=~1,tau=~s(theta,k=8,bs="cc"),
                 nu=~1,omega=~te(theta,DistanceShore,k=SP_DF,bs="cs"))

crcvm_splines_lf_he<- SDE$new(formulas = formulas,data = cropped_data,type = "RACVM_SSM",
                    response = c("y1","y2"),par0 = PAR0_RACVM,fixpar=c("mu1","mu2"),
                    other_data=list("log_sigma_obs0"=log(0.05)))

crcvm_splines_lf_he$update_map(list("log_lambda"=factor(c(1,rep(NA,2)))))
crcvm_splines_lf_he$update_lambda(c(1,10^-7,10^-13))
#fit
crcvm_splines_lf_he$fit(method="BFGS")
```

@@

crcvm_splines_lf_he$get_all_plots(show_CI="pointwise",true_smooths=list("omega"=fomega_cubic))

@@

We are clearly underestimating the magnitude of the rotation $\omega$ near $-\pi$ and $\pi$. This is indeed very similar to the result we got for the narwhal GPS data. We are well estimating $\nu$ and $\sigma_{obs}$, and we corrected the bias in $\tau$ compared to the estimate we got for the CTCRW on the same data.

@@

formulas <- list(mu1=~1,mu2=~1,tau=~1,
                 nu=~1,omega=~te(theta,DistanceShore,k=c(10,5),bs="cs"))

crcvm_splines_lf_le<- SDE$new(formulas = formulas,data = data_lf_le,type = "RACVM_SSM",
                    response = c("y1","y2"),par0 = PAR0_RACVM,fixpar=c("mu1","mu2"),
                    other_data=list("log_sigma_obs0"=log(0.03)))

crcvm_splines_lf_le$update_map(list("log_lambda"=factor(rep(NA,2))))
crcvm_splines_lf_le$update_lambda(c(10^-12,10^-12))
#fit
crcvm_splines_lf_le$fit(method="BFGS")
```

@@

crcvm_splines_lf_le$get_all_plots(show_CI="pointwise",true_smooths=list("omega"=fomega_cubic))

exp(as.list(ctcrw_lf_le$tmb_rep(),what="Estimate")$log_sigma_obs)

@@

@@

cropped_data=data_hf_he[data_hf_he$ID %in% 2:2,]

@@

crcvm_splines_hf_he<- SDE$new(formulas = formulas,data = cropped_data,type = "RACVM_SSM",
                    response = c("y1","y2"),par0 = PAR0_RACVM,fixpar=c("mu1","mu2"),
                    other_data=list("log_sigma_obs0"=log(0.05)))

#initialize smoothing penalties and re variances
new_lambda=c(10^-6,10^-4)
crcvm_splines_hf_he$update_lambda(new_lambda)

@@

plot_results_hf=crcvm_splines_hf$get_all_plots(show_CI="pointwise",true_smooths=list("omega"=fomega_cubic))

plot_results_hf

@@

For high frequency data, we are able to get $\tau$ and $\nu$ right, but we are still underestimating $\omega$.
Now we force the parametrization of the surface for $\omega$ with a new SDE model in smoothSDE (called \texttt{CRCVM_SSM}. $\omega$ is a cubic function of the angle $\theta$, with an exponential factor.

@@

cropped_data=data_hf_he[data_hf_le$ID %in% 1,]

H=array(rep(SIGMA_OBS^2*diag(2),length(data_hf$time)),dim=c(2,2,length(data_hf$time)))

@@

crcvm_splines_hf<- SDE$new(formulas = formulas,data = data_hf,type = "RACVM_SSM",
                    response = c("y1","y2"),par0 = PAR0_RACVM,fixpar=c("mu1","mu2"),
                    other_data=list("H"=H))

#initialize smoothing penalties and re variances
new_lambda=c(exp(-1),exp(-1))
crcvm_splines_hf$update_lambda(new_lambda)

crcvm_splines_hf$update_map(list(log_lambda=factor(c(NA,NA))))

#fit
crcvm_splines_hf$fit(method="BFGS")
```

@@

plot_results_hf=crcvm_splines_hf$get_all_plots(show_CI="pointwise",true_smooths=list("omega"=fomega_cubic))

plot_results_hf

## Estimation with fixed cubic function of angle

@@

```{r}
H=array(rep(0.03^2*diag(2),length(data_lf_he$time)),dim=c(2,2,length(data_lf_he$time)))
# Estimate from simulated data ------------------

@@

crcvm_cubic_lf_he<- SDE$new(formulas = formulas,data = data_lf_he[data_lf_he$DistanceShore>0.01,],type = "CRCVM_SSM",
                    response = c("y1","y2"),par0 = c(1,1,1,5,0.5,0.8,0.3,pi/8),
                    fixpar=c("a","b","D0","D1","sigma_D","sigma_theta"),
                    other_data=list("log_sigma_obs0"=log(0.02)))
#fit
crcvm_cubic_lf_he$fit(method="BFGS")
```

@@

```{r}
H=array(rep(0.007^2*diag(2),length(data_lf_le$time)),dim=c(2,2,length(data_lf_le$time)))
# Estimate from simulated data ------------------

@@

crcvm_cubic_lf_le<- SDE$new(formulas = formulas,data = data_lf_le,type = "CRCVM_SSM",
                    response = c("y1","y2"),par0 = c(1.5,4,1,5,1,0.8,0.2,pi/8),fixpar=c("a","b","D0","D1","sigma_D","sigma_theta"),
                    other_data=list("log_sigma_obs0"=log(0.03)))
#fit
crcvm_cubic_lf_le$fit(method="BFGS")
```
```{r}

@@

@@

@@

crcvm_cubic_hf_he<- SDE$new(formulas = formulas,data = data_hf_he[data,type = "CRCVM_SSM",
                    response = c("y1","y2"),par0 =c(1,1,1,5,0.5,0.8,0.3,pi/8),
                    fixpar=c("a","b","D0","D1","sigma_D","sigma_theta"),
                    other_data=list("log_sigma_obs0"=log(0.05)))

#fit
crcvm_cubic_hf_he$fit(method="BFGS")
```

```{r}
crcvm_cubic_hf_he$get_all_plots()
exp(as.list(crcvm_cubic_hf_he$tmb_rep(),what="Estimate")$log_sigma_obs)
```
### High frequency low error

crcvm_cubic_hf_le<- SDE$new(formulas = formulas,data = data_hf_le[data_hf_le$ID %in% 4:5,],type = "CRCVM_SSM",
                    response = c("y1","y2"),par0 = c(1,1,40,5,0.15,0.8,0.2,pi/8),fixpar=c("a","b","D0","D1","sigma_D","sigma_theta"),
                    other_data=list("log_sigma_obs0"=log(0.03)))

@@

```{r}
# Definition of parameters tau and nu ----------------
tau_re=rnorm(N_ID,mean=0,sd=SIGMA_TAU)
nu_re=rnorm(N_ID,mean=0,sd=SIGMA_NU)
true_log_nu=nu_re+log(NU_0)

ftau_gaussian=function(cov_data,sigma_bump=0.5,sigma_dip=0.5,tau0=TAU_0,dtau_bump=1.5,dtau_dip=0) {
  #cov_data : dataframe of covariates with column for angle theta
  #sigma_bump : width of the bumps at -pi/2 and pi/2
  #sigma_dip : width of the dip at 0
  #tau0 : base value of tau
  #dtau_bump : height of the bump
  #dtau_dip : depth of the dip

  theta=cov_data$theta
  bump1 <- dtau_bump*exp(-((theta - pi/2)^2) / (2 * sigma_bump^2))
  bump2 <- dtau_bump*exp(-((theta + pi/2)^2) / (2 * sigma_bump^2))
  dip<- dtau_dip*exp(-((theta)^2) / (2 * sigma_dip^2))
  tau=tau0+bump1+bump2-dip

  return (tau)
}

>>>>

```{r}
library(here)
library(sf)
library(smoothSDE)
library(doParallel)
library(ggplot2)
library(dplyr)
library(mgcv)
library(plotly)

@@

```{r hyperparameters}
N_ID=6

TMAX=12
DELTA=1/60/6

SP_DF=c(5,5)

TAU_0=1.3
NU_0=4.6

DMIN=0.5
DMAX=1

PAR0_RACVM=c(0,0,1,1,0)

SIGMA_OBS_LOW=0.01
SIGMA_OBS_HIGH=0.04

BY_LF=5
BY_HF=1

D_LOW=0.05
D_UP=3

A=1
D0=0.3
D1=1.2
SIGMA_THETA=pi/6
SIGMA_D=1
B=1

@@

seed=42
set.seed(seed)
# Set the path to the directory containing the data
greenland_data_path <- here("Data","preprocessed_data","greenland")

@@

shrunk_border<-st_read(file.path(greenland_data_path,"shrunk_updated_scoresby_sound_utm.shp"))
shrunk_border <- st_transform(border, crs = "+init=EPSG:32626 +units=km")
v0=c(0,0)

@@

i=1
while (i<=N_ID) {
  #choose location uniformly in the map
  x=c(runif(1,min=430,max=500),runif(1,min=7760,max=7900))
  if (!(is_in_land(st_point(x),border))) {
    p=nearest_boundary_point(st_point(x),border)
    Dshore=(x[1]-p[1])^2+(x[2]-p[2])^2
    #keep it as initial position if it is at least 50 metres away from the shore
    if (Dshore>DMIN && Dshore<DMAX) {
      x0[i,]=x
      i=i+1
    }
  }
}

@@

fomega_cubic=function(cov_data,a,b,D0,D1,sigma_D,sigma_theta){
  Dshore=cov_data$DistanceShore
  theta=cov_data$theta
  if (is.null(Dshore)){
    Dshore=1/cov_data$Ishore
  }

@@

fomega=function(cov_data) {
  fomega_cubic(cov_data,A,B,D0,D1,SIGMA_D,SIGMA_THETA)
}

@@

  res=sim_CRCVM(ftau=ftau_constant,fomega=fomega,fnu=fnu_constant,
                      log_sigma_obs=NULL,v0=v0,x0=x0[i,],times=times,
                land=shrunk_border,verbose=FALSE,save_omega = TRUE)

  data_sim=res$sim
  data_sim$ID=factor(rep(i,length(data_sim$y1)))
  data_shore=res$shore

  data_sim=cbind(data_sim,data_shore)
  data_sim
}

@@

```{r}
# add noise
low_noise=rmvn(n_obs,rep(0,2),diag(rep(SIGMA_OBS_LOW^2,2)))
high_noise=rmvn(n_obs,rep(0,2),diag(rep(SIGMA_OBS_HIGH^2,2)))
observed_data_le=data_sub[,c("y1","y2","time","ID")]
observed_data_le[,c("y1","y2")]=data_sub[,c("y1","y2")]+low_noise
observed_data_he=data_sub[,c("y1","y2","time","ID")]
observed_data_he[,c("y1","y2")]=data_sub[,c("y1","y2")]+high_noise

#subsample
data_lf_he=observed_data_he[seq(1,length(data_sub$time),by=BY_LF),]
data_lf_le=observed_data_le[seq(1,length(data_sub$time),by=BY_LF),]
data_hf_le=observed_data_le[seq(1,length(data_sub$time),by=BY_HF),]
data_hf_he=observed_data_he[seq(1,length(data_sub$time),by=BY_HF),]

@@

plot=ggplot()+geom_sf(data=border$geometry,fill="grey")+
  coord_sf(datum=st_crs("+init=EPSG:32626 +units=km"))+
  geom_point(data=data_sub,mapping=aes(y1,y2,col=ID),size=0.1,alpha=0.2,shape=2)+
  geom_point(data=data_lf_he,mapping=aes(y1,y2,col=ID),size=0.1)+
  geom_path(data=data_lf_he,mapping=aes(y1,y2,col=ID),size=0.1)+
  geom_point(data = data%>% filter(!duplicated(ID)),
             aes(x = y1, y = y2), shape = 3, size = 4, col = "red")+
  xlab("x") + ylab("y")

ggsave(filename="plot_trajectories.pdf",plot=plot,width=10,height=5)

@@

  # Function to calculate rolling average
rolling_average <- function(data, window) {
  n <- nrow(data)
  result <- matrix(NA, n, ncol(data))  # Initialize result matrix with NA
  colnames(result) <- colnames(data)

  for (i in seq_len(n)) {
    # Define the range for the current window
    start <- max(1, i - floor(window / 2))
    end <- min(n, i + floor(window / 2))
    result[i, ] <- colMeans(data[start:end, , drop = FALSE])  # Compute column means
  }
  return(result)
}

add_covs_parallel <- function(data,add="all",window=1,n_cores = parallel::detectCores() - 1) {
  # Setup parallel backend
  cl <- makeCluster(n_cores)
  registerDoParallel(cl)

  # Split data by ID
  ids <- unique(data$ID)

  # Process each ID in parallel
  results <- foreach(id = ids, .combine = rbind, .packages = c("sf"),
                     .export = c("nearest_boundary_points", "is_in_land", "signed_angle","border",
                                 "D_LOW","D_UP","rolling_average")  ) %dopar% {
    # Filter data for the current ID
    sub_data <- data[data$ID == id, ]
    n_sub <- nrow(sub_data)

    # Time steps
    dtimes <- sub_data[2:n_sub, "time"] - sub_data[1:(n_sub - 1), "time"]

    # Step lengths
    dx <- sub_data[2:n_sub, "y1"] - sub_data[1:(n_sub - 1), "y1"]
    dy <- sub_data[2:n_sub, "y2"] - sub_data[1:(n_sub - 1), "y2"]

    # Empirical velocity
    vexp_df <- cbind(dx / dtimes, dy / dtimes)

    vexp_avg <- rolling_average(vexp_df, window)

    # Nearest points on shore
    sub_data <- cbind(sub_data, nearest_boundary_points(as.matrix(sub_data[, c("y1", "y2")]), border))

    # Normal vectors
    normal <- as.matrix(sub_data[2:n_sub, c("y1", "y2")] - sub_data[2:n_sub, c("p1", "p2")])

    # Angle between velocity and normal vector
    theta_coast <- signed_angle(normal, vexp_avg)
    theta_coast <- c(theta_coast, 1)  # Adjust length

    # Initialize DistanceShore
    Dshore <- rep(0, n_sub)

    for (i in 1:n_sub) {
      y1 <- sub_data[i, "y1"]
      y2 <- sub_data[i, "y2"]
      if (!is_in_land(st_point(c(y1, y2)), border)) {
        p1 <- sub_data[i, "p1"]
        p2 <- sub_data[i, "p2"]
        Dshore[i] <- sqrt((y1 - p1)^2 + (y2 - p2)^2)
      }
    }

    Dshore_avg=as.vector(rolling_average(matrix(Dshore,ncol=1),window))

    # Calculate Ishore
    Ishore_avg <- ifelse(Dshore_avg < D_LOW, 1 / D_LOW, ifelse(Dshore_avg > D_UP, 0, 1 / Dshore_avg))

    # Add columns to sub_data
    if (add=="all") {
      sub_data$theta <- theta_coast
      sub_data$DistanceShore <- Dshore_avg
      sub_data$Ishore <- Ishore_avg
    }
    else if (add=="theta") {
      sub_data$theta <- theta_coast
    }
    else if (add=="DistanceShore") {
      sub_data$DistanceShore <- Dshore
      sub_data$Ishore <- Ishore
        }
  else {
    stop("Options for argument add are \"all\",\"DistanceShore\" and \"theta\"")
  }

    return(sub_data)
  }

  # Stop the cluster
  stopCluster(cl)

  return(results)
}

@@

```{r}
#colnames(data)[8]="DistanceShore"
data_lf_he=add_covs_parallel(data_lf_he,window=3)
data_hf_le=add_covs_parallel(data_hf_le,window=3)
data_lf_le=add_covs_parallel(data_lf_le,window=3)
data_hf_he=add_covs_parallel(data_hf_he,window=3)
```

@@

ggplot(data_lf_he)+geom_histogram(aes(x=DistanceShore),bins=30)+facet_wrap(~ID)+theme_minimal()
quantile(data_lf_he$DistanceShore)
```

@@

ctcrw_lf_he<- SDE$new(formulas = formulas,data =data_lf_he[data_lf_he$ID %in% c(1,2,4,5),],type = "CTCRW",
                    response = c("y1","y2"),par0 = c(0,0,1,1),fixpar=c("mu1","mu2"),
                    other_data=list("log_sigma_obs0"=log(0.05)))

@@

formulas <- list(mu1=~1,mu2=~1,tau=~1,nu=~1,omega=~1)

ctcrw_lf_le<- SDE$new(formulas = formulas,data =data_lf_le[data_lf_le$ID %in% c(1:6),],type = "RACVM_SSM",
                    response = c("y1","y2"),par0 = c(0,0,1,1,0),fixpar=c("mu1","mu2"),
                    other_data=list("log_sigma_obs0"=log(0.03)))

@@

#points on the surface perturbed by gaussian noise
f <- fomega(samples)
y <- f+0.1*rnorm(n)

@@

formulas <- list(mu1=~1,mu2=~1,tau=~1,
                 nu=~1,omega=~te(theta,DistanceShore,k=SP_DF,bs=c("cs")))

crcvm_splines_lf_he<- SDE$new(formulas = formulas,data = data_lf_he,type = "RACVM_SSM",
                    response = c("y1","y2"),par0 = PAR0_RACVM,fixpar=c("mu1","mu2"),
                    other_data=list("log_sigma_obs0"=log(0.05)))

crcvm_splines_lf_he$update_map(list("log_lambda"=factor(c(rep(NA,2)))))
crcvm_splines_lf_he$update_lambda(m1$sp)
#fit
crcvm_splines_lf_he$fit(method="BFGS")
```

@@

crcvm_splines_lf_he$get_all_plots(show_CI="pointwise",true_smooths=list("omega"=fomega))

@@

We are clearly underestimating the magnitude of the rotation $\omega$ near $-\pi$ and $\pi$. This is indeed very similar to the result we got for the narwhal GPS data. We are well estimating $\nu$ and $\sigma_{obs}$, and we corrected the bias in $\tau$ compared to the estimate we got for the CTCRW on the same data.
We can try to fit the true parametric function on the estimated spline surface.

estimate_parametric_omega=function(model) {

    data=model$data()

    Dshore=seq(from=quantile(data$DistanceShore,0.5),to=quantile(data$DistanceShore,0.95),length.out=30)
    theta=seq(from=-pi,to=pi,length.out=30)
    grid<- as.data.frame(expand.grid(Dshore,theta))
    colnames(grid) <- c("DistanceShore","theta")
    grid$ID=1

    #get model matrices
    mats=model$make_mat(new_data=grid)
    X_fe=mats$X_fe
    X_re=mats$X_re

    par_mat=model$par(new_data=grid,X_fe=X_fe,X_re=X_re)

    #matrix of values for surface plot
    z=matrix(par_mat[,"omega"],30,30)

    data=grid
    data$z=as.vector(z)

    # Create the 3D surface plot using plot_ly
    plot_splines<- plot_ly(x = ~theta, y = ~Dshore, z = ~z) %>%
      add_surface() %>%
      layout(title = "Predicted spline surface for omega",
             scene = list(yaxis = list(title = "Distance to shore"),
               xaxis = list(title = "theta")))

    fit_optim <- optim(par = c(0.1,2),
                       fn = function(params) {

                        # Compute predicted z values
                        z_pred <- fomega_cubic(data,A,B,params[1],params[2],SIGMA_D,SIGMA_THETA)

                         # Return sum of squared residuals
                         sum((data$z - z_pred)^2)
                       })

    est_par=fit_optim$par

    # Compute the predicted values using the starting parameters
    z_pred_starting <- fomega_cubic(grid,A,B,est_par[1],est_par[1],SIGMA_D,
                                    SIGMA_THETA)

    # Reshape z_pred_starting to a matrix for plotting
    z_pred_matrix <- matrix(z_pred_starting, nrow = length(Dshore), ncol = length(theta))

    # Create the 3D surface plot using plot_ly
    plot_parametric <- plot_ly(x = ~theta, y = ~Dshore, z = ~z_pred_matrix) %>%
      add_surface() %>%
      layout(title = "Estimated parametric surface for omega",
             scene = list(
               yaxis = list(title = "Distance to shore"),
               xaxis = list(title = "theta"),
               zaxis = list(title = "Predicted z")
             ))

    return (list("est_par"=est_par,"splines_surface"=plot_splines,"parametric_surface"=plot_parametric))

estimate_parametric_omega(crcvm_splines_lf_he)

@@

crcvm_splines_lf_le<- SDE$new(formulas = formulas,data = data_lf_le[data_lf_le$ID %in% c(1:6),],type = "RACVM_SSM",
                    response = c("y1","y2"),par0 = PAR0_RACVM,fixpar=c("mu1","mu2"),
                    other_data=list("log_sigma_obs0"=log(0.05)))

crcvm_splines_lf_le$update_map(list("log_lambda"=factor(rep(NA,2))))
crcvm_splines_lf_le$update_lambda(m1$sp)
#fit
crcvm_splines_lf_le$fit(method="BFGS")
```

@@

crcvm_splines_lf_le$get_all_plots(show_CI="pointwise",true_smooths=list("omega"=fomega))

exp(as.list(crcvm_lf_le$tmb_rep(),what="Estimate")$log_sigma_obs)

@@

```{r}
estimate_parametric_omega(crcvm_splines_lf_le)
```

@@

@@

crcvm_splines_hf_he<- SDE$new(formulas = formulas,data = data_hf_he[data_hf_he$ID %in% 1:2,],type = "RACVM_SSM",
                    response = c("y1","y2"),par0 = PAR0_RACVM,fixpar=c("mu1","mu2"),
                    other_data=list("log_sigma_obs0"=log(0.05)))

#initialize smoothing penalties and re variances
new_lambda=c(m1$sp)
crcvm_splines_hf_he$update_lambda(new_lambda)

@@

crcvm_splines_hf_he$get_all_plots(show_CI="pointwise",true_smooths=list("omega"=fomega))

@@

estimate_parametric_omega(crcvm_splines_hf_he)

@@

@@

crcvm_splines_hf_le<- SDE$new(formulas = formulas,data = data_hf_le[data_hf_le$ID %in% 1:1,],type = "RACVM_SSM",
                    response = c("y1","y2"),par0 = PAR0_RACVM,fixpar=c("mu1","mu2"),
                    other_data=list("log_sigma_obs0"=log(0.05)))

#initialize smoothing penalties and re variances
new_lambda=c(m1$sp)
crcvm_splines_hf_le$update_lambda(new_lambda)

crcvm_splines_hf_le$update_map(list(log_lambda=factor(c(NA,NA))))

#fit
crcvm_splines_hf_le$fit(method="BFGS")
```

```{r}
crcvm_splines_hf_le$get_all_plots(show_CI="pointwise",true_smooths=list("omega"=fomega))

@@

estimate_parametric_omega(crcvm_splines_hf_le)
```

## Estimation with fixed parametric omega function

@@

```{r}
H=array(rep(SIGMA_OBS_HIGH^2*diag(2),length(data_lf_he$time)),dim=c(2,2,length(data_lf_he$time)))
# Estimate from simulated data ------------------

@@

crcvm_cubic_lf_he<- SDE$new(formulas = formulas,
                            data = data_lf_he,type = "CRCVM_SSM",
                    response = c("y1","y2"),par0 = c(1,1,A,B,D0,D1,SIGMA_D,SIGMA_THETA),
                    fixpar=c("a","b","D0","D1","sigma_D","sigma_theta"),
                    other_data=list("H"=H))

#fit
crcvm_cubic_lf_he$fit(method="BFGS")
```

@@

```{r}
H=array(rep(SIGMA_OBS_LOW^2*diag(2),length(data_lf_le$time)),dim=c(2,2,length(data_lf_le$time)))
# Estimate from simulated data ------------------

@@

crcvm_cubic_lf_le<- SDE$new(formulas = formulas,data = data_lf_le[data_lf_le$ID %in% c(1:4),],
                              type = "CRCVM_SSM",
                    response = c("y1","y2"),par0 = c(1,1,A,B,D0,D1,SIGMA_D,SIGMA_THETA),fixpar=c("a","b","D0","D1","sigma_D","sigma_theta"),
                    other_data=list("H"=H))

crcvm_cubic_lf_le$fit(method="BFGS")
```
```{r}

@@

```{r}
H=array(rep(SIGMA_OBS_HIGH^2*diag(2),length(data_hf_he$time)),dim=c(2,2,length(data_hf_he$time)))

crcvm_cubic_hf_he<- SDE$new(formulas = formulas,data = data_hf_he[data_hf_he$DistanceShore>0.01 & data_hf_he$ID %in% c(1:6),],type = "CRCVM_SSM",
                    response = c("y1","y2"),par0 =c(1,1,A,B,D0,D1,SIGMA_D,SIGMA_THETA),
                    fixpar=c("a","b","D0","D1","sigma_D","sigma_theta"),
                    other_data=list("H"=H))

crcvm_cubic_hf_he$fit(method="BFGS")
```

```{r}
crcvm_cubic_hf_he$get_all_plots()
exp(as.list(crcvm_cubic_hf_he$tmb_rep(),what="Estimate")$log_sigma_obs)
```

@@

H=array(rep(SIGMA_OBS_LOW^2*diag(2),length(data_hf_le$time)),dim=c(2,2,length(data_hf_le$time)))

@@

crcvm_cubic_hf_le<- SDE$new(formulas = formulas,data = data_hf_le[data_hf_le$ID %in% c(1:4),],
                            type = "CRCVM_SSM",
                    response = c("y1","y2"),par0 = c(1,1,A,B,D0,D1,SIGMA_D,SIGMA_THETA),
                    fixpar=c("a","b","D0","D1","sigma_D","sigma_theta"),
                    other_data=list("H"=H))

@@

```{r}
ggplot()+geom_point(data=data_hf_le,aes(time,theta),col="blue",alpha=0.5,size=0.1)+
  geom_point(data=data,aes(time,theta),col="red",alpha=0.2,size=0.1)+facet_wrap(~ID)

```
```{r}
data_hf_le$omega=fomega(data_hf_le[,c("theta","DistanceShore")])

ggplot()+geom_point(data=data_hf_le,aes(time,omega),col="blue",alpha=0.5,size=0.1)+
  geom_point(data=data,aes(time,omega),col="red",alpha=0.2,size=0.1)+facet_wrap(~ID)
```
